A quick user guide to the\\ \scine-\qcm\ software suite for \mol

### 2023-05-08 dmrg/doc/qcmaquis_manual.tex
Version 3.1.3

% -- Loads packages --
\usepackage[nott]{kpfonts}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{booktabs}
\usepackage[bf,singlelinecheck=off]{caption}
\usepackage[T1]{fontenc}
\usepackage{framed}
\usepackage{graphicx}
\usepackage[colorlinks=true]{hyperref} % web links
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{longtable}
\usepackage{multirow}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{placeins}
\usepackage{xcolor}

@@

\lstdefinelanguage{qcmaquis}
{
%
    keywords=[1]{L,nelec,spin,irrep,nsweeps,max_bond_dimension,integral_file,chkpfile,resultfile,init_state,LATTICE,lattice_library,MODEL,model_library},
	morekeywords=[2]{conv_thresh,integral_cutoff,truncation_initial,truncation_final,symmetry,MEASURE},
	keywordstyle=[1]\color{green!20!red},
	keywordstyle=[2]\color{green!20!blue},
	commentstyle=\color{gray},
	sensitive=false,
	morecomment=[l]{//},
	morecomment=[s]{/*}{*/},
}

@@

\section{\qcm{} Standalone Version}

As already anticipated, \qcm\ can be used also as a standalone code.
The standalone version of \qcm\ is used especially in two cases: (i) for electronic-structure calculations relying on integrals that are calculated with programs other than \mol{} and (ii) for DMRG calculations beyond time-independent electronic-structure simulations.
In the present section, we focus on the former application field, while the latter will be discussed in the following Sections. \\
Section~\ref{sec:install-qcm-stand}\ describes in details the installation process for a standalone version of \qcm.

\begin{framed}
 \noindent Note: If you have installed \qcm{} with the \qcm{}-\mol{} interface, you may skip this Section.
\end{framed}

\subsection{Installation}
\label{sec:install-qcm-stand}

\noindent We describe in the following how to build and install the standalone version of the \qcm\ software.
The installation of \qcm\ has been tested for different operating systems and compiler/math libraries environments.
Their list can be found in Section \ref{subsec:supportedOS}.
While other combinations might work equally well they are \emph{not} officially supported.

The following libraries are required and need to be installed manually by the user (but are usually provided by the system package manager)

\begin{itemize}
 \item GSL
 \item HDF5
 \item SZIP
 \item Boost
\end{itemize}

Additionally, \qcm{} distribution includes a customised version of the ALPS 2.3.0 library, which will be compiled automatically.

\subsubsection{Download \qcm}
\label{subsubsec:download-stand}

As mentioned in Section~\ref{sec:reg}, the \qcm{} standalone version may be found in \texttt{\url{https://github.com/qcscine/qcmaquis}}.

\subsubsection{Setting up a build folder}

Clone the GitHub repository into a new directory \texttt{\qcmsrc}. Create a build folder \texttt{\qcmbuild} -- note that this folder \emph{does not} necessarily have to be a subfolder of \qcmsrc\ -- and change to this new folder:

\begin{verbatim}
  mkdir /path/to/my-QCMaquis-build && cd /path/to/my-QCMaquis-build
\end{verbatim}

\subsubsection{Configuration}
\label{subsubsec:configure-stand}

\qcm{} supports operating systems, C++ compilers and math libraries as listed in Section~\ref{subsec:supportedOS}. 
Below we will show the configuration steps for the most popular compiler suites GNU (Section \ref{sec:gnu-conf-stand}) and Intel (Section \ref{sec:intel-conf-stand}), respectively.
How to setup and use a shared-memory OMP installation of \qcm\ is described in Section \ref{sec:parallel-stand}.

\paragraph{Configuration with the GNU compiler suite}
\label{sec:gnu-conf-stand}$\;$\\

\noindent To configure \qcm\ with the GNU compiler suite type

\begin{verbatim}
  CC=gcc CXX=g++ cmake /path/to/my-QCMaquis-src
\end{verbatim}
%
where we assumed that the Intel Math Kernel Library (MKL) is available (recommended option).
If the MKL libraries are not available, \qcm\ will search for other suitable math libraries installed on the operating system. If none are found the configuration step will stop with an appropriate message.

% \paragraph{Configuration with the Intel compiler suite}\label{sec:intel-conf-stand}$\;$\\
%
% \noindent To configure \qcm\ with the Intel compiler suite including MKL type
% \begin{verbatim}
% FC=ifort CC=icc CXX=icpc cmake -DQCM_standalone=ON /path/to/my-QCMaquis-src
% \end{verbatim}

\paragraph{Shared-memory OMP parallel configuration}
\label{sec:parallel-stand}$\;$\\

By default \qcm\ is built as shared-memory OMP parallelized version which should work with either compiler suite, GNU or Intel.

In order to exploit the shared-memory OMP parallelization of \qcm\ the user is strongly encouraged to set at runtime the environment variable

\begin{verbatim}
  export OMP_NUM_THREADS=XX
\end{verbatim}
%
where \texttt{XX}\ specifies the number of shared-memory cores to be used.
The default (\emph{depending on the operating system!!!}) is to use all available cores.

\subsubsection{Building and installation}
\label{subsubsec:build-stand}

After a successful configuration, type
\begin{verbatim}
  make
\end{verbatim}
%
or

\begin{verbatim}
  make -j8
\end{verbatim}
%
to compile \qcm\ (in parallel on 8 cores) and install all its components in the build folder \texttt{\qcmbuild}.

\subsection{Keywords and Options}\label{sec:qcmaquis-kw}

In the following we describe (i) compulsory keywords (Section \ref{sec:compul-qcm}), (ii) optional keywords (Section \ref{sec:optional-qcm}) for \qcm\ DMRG calculations as well as (iii) keywords for property calculations (Section \ref{sec:prop-qcm}).
Most of the \qcm\ keywords have default settings that guarantee convergence in the general case and are inserted automatically
by the host program (\hostp\ in this case). 
A reasonable choice of default values for optional keywords is given in our
example \qcm\ input file in Section~\ref{sec:input-qcm}.

\begin{framed}
  \noindent\textbf{Note:}\ \textsc{Maquis} has many features beyond quantum chemistry, e.g. related to solid state physics.
  Some keywords listed in the example file in Section \ref{sec:input-qcm} (e.\,g.\ \kwd{LATTICE}, \kwd{lattice\_library} etc.)\ are therefore explained only briefly in the following and are not to be changed from the default values (see Listing~\ref{lst:qcm}), if a quantum chemical calculation is desired.
\end{framed}

\subsubsection{Compulsory keywords}
\label{sec:compul-qcm}

The keywords in Tables \ref{tab:standard_kw} and \ref{tab:standard_kw_auto} have to be set for every DMRG calculation since they may crucially affect the accuracy of the final result.
Their choice depends for example on the molecule under consideration (i.\,e. whether you expect strong static electron correlation and/or dynamical correlation to play a major role), the nature of the reference orbitals (Hartree-Fock orbitals, natural orbitals of some kind, \ldots), the active space size, and many other aspects.

% \noindent Some of the compulsory keywords listed in Table \ref{standard_kw}\ are indeed automatically set if you run a \qcm\ DMRG calculation through the \qcm\ host program driver which is the case for \mol-\qcm\ DMRG
% calculations. In this case skip the upper part of Table \ref{standard_kw}\ and proceed immediately to the lower part
% marked by ``{\color{red}{Keywords \emph{NOT} set by the host program \mol}}".

\begin{table}
  \caption{\label{tab:standard_kw_auto}Compulsory \qcm{} keywords automatically set by the \qcm{}-\mol{} interface.}
  \begin{tabular}{p{3.6cm}l}
  \toprule
keyword & description \\
  \midrule

\texttt{nelec} & Total number of electrons. \\[\tableoptionskip]
\multirow{3}{*}{\texttt{irrep}} & \multirow{3}{12cm}{Irreducible representation of the point group symmetry of the target state. \textbf{Note:} Counting starts with 0 which has to be the totally symmetric representation.} \\
& \\
& \\[\tableoptionskip]
\multirow{2}{*}{\texttt{spin}} & \multirow{2}{12cm}{Total spin ($2\times S$) of the target state, for example:\\
\texttt{spin=0} (singlet), \texttt{spin=1} (doublet), \texttt{spin=2} (triplet), \ldots}\\
& \\[\tableoptionskip]
\texttt{L} & Length of lattice = number of orbitals in the active space. \\[\tableoptionskip]
\multirow{2}{*}{\texttt{integral\_file}} & \multirow{2}{12cm}{Path and filename of the integral file, for example\\
\texttt{integral\_file = /path/to/file/FCIDUMP}} \\
& \\[\tableoptionskip]
\texttt{chkpfile} & Path and name of the folder in which the MPS is stored. \\[\tableoptionskip]
\texttt{resultfile} & Path and filename of the result file. \\[\tableoptionskip]
\multirow{3}{*}{\texttt{n\_ortho\_states}} &\multirow{3}{12cm}{If an excited state calculation is desired, the number of states the current wave function is to be orthogonalized against shall be specified here.} \\
& \\
& \\[\tableoptionskip]
\multirow{3}{*}{\texttt{ortho\_states}} & \multirow{3}{12cm}{Path(s) and filename(s) of the MPS checkpoint file(s) that store the lower lying states to which the current MPS shall be orthogonal to.} \\
& \\
& \\[\tableoptionskip]
\multirow{9}{*}{\texttt{init\_state}} & \multirow{9}{12cm}{Possible options are \texttt{default}, \texttt{thin} and \texttt{hf}.
The \texttt{default} and \texttt{thin} initializations fill the initial MPS with random numbers, the difference being
that a singular value decomposition reduces the bond dimension to \texttt{init\_bond\_dimension} in the case of
\texttt{thin}. Usage of \texttt{hf} generates an MPS consisting of only the determinant defined on the \texttt{hf\_occ}
card. Note that the CI-DEAS procedure \cite{Legeza2003,fiedler,interface} (as invoked by \texttt{dmrginit.py}, see Section
\ref{sec:pytools-qcm}) behaves like a restart from the newly generated CI-DEAS MPS.\\
% {\textbf{default in \mol: \texttt{init\_state = "default"}}}
}\\
& \\
& \\
& \\
& \\
& \\
& \\
& \\
& \\[\tableoptionskip]
\multirow{7}{*}{\begin{minipage}{5cm}
                  \texttt{%
                  LATTICE\\
                  lattice\_library\\
                  MODEL\\
                  model\_library\\}
                \end{minipage}}
 & \multirow{3}{11cm}{These options set the Hamiltonian and the lattice type in \qcm{} calculations. For a quantum chemical calculation, these must be set to (see also Listing~\ref{lst:qcm})}\\
 & \\
 & \\
  & \texttt{LATTICE = "orbitals"}\\
  & \texttt{lattice\_library = "coded"}\\
  & \texttt{MODEL = "quantum\_chemistry"}\\
  & \texttt{model\_library = "coded"}\\
\bottomrule
\end{tabular}
\end{table}
\FloatBarrier

\begin{framed}
  \noindent\textbf{Note:}\ We strongly encourage any new user of \qcm\ to carefully read first Ref.\ \citenum{Keller_CHIMIA_Determining_2014}\ which gives a detailed introduction to
  DMRG calculations in quantum chemistry. Further quantum chemical DMRG sample calculations starting from different computational setups are discussed for example in Refs. \cite{interface,ru-dmrg,fde-dmrg,srdft-dmrg}.
\end{framed}

\subsubsection{Optional keywords}
\label{sec:optional-qcm}

The keywords summarized in Table \ref{tab:advanced_kw} may be exploited by the more experienced user but can be safely
ignored by those who just want to get started. They may affect the convergence and accuracy of the final result, though.
For the inexperienced user however, we advise to not change these settings and accept the default values provided by \qcm.

% \vspace{3ex}

\begin{longtable}{ll}
\caption{Optional keywords for \qcm\ calculations.}\label{tab:advanced_kw}
\\
  \toprule
keyword & description \\
  \midrule
\endfirsthead

\multicolumn{2}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\toprule
keyword & description \\
\midrule
\endhead

\multirow{6}{*}{\texttt{orbital\_order}} & \multirow{6}{12cm}{Manual ordering of the orbitals along the one dimensional lattice. The order has to be entered as a string of comma separated orbital numbers. We recommend the Fiedler ordering based on the mutual information \cite{interface,fiedler} which can be obtained by means of the python script \texttt{fiedler.py} (see Section \ref{sec:pytools-qcm}).\\
	\textbf{default}: \texttt{orbital\_order = "1,2,3,4,5,6,\ldots"}} \\
& \\[\tableoptionskip]
\multirow{4}{*}{\texttt{sweep\_bond\_dimensions}} & \multirow{4}{12cm}{Sets \texttt{max\_bond\_dimension}\ for each sweep separately.\\
	\emph{Note}: Replaces \texttt{max\_bond\_dimension} which does \emph{NOT} need to be specified in this case.\\
	Example (\texttt{nsweeps=3}): \texttt{sweep\_bond\_dimensions}=``300,400,500"} \\
& \\[\tableoptionskip]
\multirow{2}{*}{\texttt{init\_bond\_dimension}} & \multirow{2}{12cm}{Adjusts the maximal bond dimension of the MPS produced by the
CI-DEAS procedure \cite{Legeza2003,fiedler,interface}.} \\
& \\[\tableoptionskip]
\multirow{7}{*}{\texttt{conv\_thresh}} &\multirow{7}{12cm}{Sets the energy convergence threshold (in Hartree). If the lowest energy from the previous sweep differs from the lowest energy of the current sweep by less than \texttt{conv\_thresh}, the DMRG calculation stops.\\
		\emph{Note}: Requires to set also \texttt{truncation\_final} and \texttt{ietl\_jcd\_tol}.\\
	{Numerical format}: \texttt{xe-y} with \texttt{x}\ and \texttt{y} being integers. \\
	Example: \texttt{conv\_thresh = 1e-6}.}\\
& \\[\tableoptionskip]
\multirow{3}{*}{\texttt{ietl\_jcd\_tol}} & \multirow{3}{12cm}{Convergence threshold for the Jacobi-Davidson diagonalization.\\
	{Numerical format}: \texttt{xe-y} with \texttt{x}\ and \texttt{y} being integers. \\
	Example: \texttt{ietl\_jcd\_tol = 1e-6}.} \\
& \\[\tableoptionskip]
\multirow{2}{*}{\texttt{ietl\_jcd\_maxiter}} & \multirow{2}{12cm}{Maximum number of iterations in the Jacobi-Davidson diagonalization.} \\
& \\ \pagebreak
\multirow{6}{*}{\texttt{truncation\_initial}} & \multirow{6}{12cm}{If during the \texttt{ngrowsweeps}, the sum of the discarded singular values for $m$ retained states is lower than the value defined here, more block states will be discarded until the discarded sum increases to \texttt{truncation\_initial}.\\
		{Numerical format}: \texttt{xe-y} with \texttt{x}\ and \texttt{y} being integers. \\
		Example: \texttt{truncation\_initial = 1e-6}.}\\
& \\[\tableoptionskip]
\multirow{6}{*}{\texttt{truncation\_final}} & \multirow{6}{12cm}{If during the \texttt{nmainsweeps}, the sum of the discarded singular values for $m$ retained states is lower than the value defined here, more block states will be discarded until the discarded sum increases to \texttt{truncation\_final}.\\
	{Numerical format}: \texttt{xe-y} with \texttt{x}\ and \texttt{y} being integers. \\
	Example: \texttt{truncation\_final = 1e-6}.}\\
& \\[\tableoptionskip]
\multirow{2}{*}{\texttt{measure\_each}} &\multirow{2}{12cm}{Tells the program to compute the expectation values every\\$2\ \times$
	\texttt{measure\_each} sweeps.} \\
& \\[\tableoptionskip]
\multirow{10}{*}{\texttt{symmetry}} & \multirow{10}{11cm}{Defines the total symmetry group. Default is the combined spin- (SU2) and point
symmetry group (PG) \texttt{su2u1pg}, where the \texttt{pg}-suffix should be omitted for better performance if the molecule is C1-symmetric.
For test purposes, it is possible to switch off spin-adaptation, again with or without point group symmetry: \texttt{2u1(pg)} .
In the latter case the keywords \texttt{spin} and \texttt{nelec} (see Section \ref{sec:compul-qcm}) have no meaning.
Instead \texttt{u1\_total\_charge1} and \texttt{u1\_total\_charge2} corresponding to the number of up and down electrons have to be specified.} \\
& \\[\tableoptionskip]
\multirow{2}{*}{\texttt{chkp\_each}} & \multirow{2}{11cm}{Tells the program to update the checkpoint file every\\ $2\ \times$
\texttt{chkp\_each} sweeps.} \\
& \\[\tableoptionskip]
\multirow{9}{*}{\texttt{hf\_occ}} & \multirow{9}{11cm}{Occupation of the starting orbitals (e.g. Hartree-Fock occupation) to be
entered as a comma separated string of occupation aliases. The aliases are defined as follows: 4~=~full, 3~=~up,
2~=~down, 1~=~empty. This information has to be entered in case of \texttt{hf} as \texttt{init\_state} and for the
CI-DEAS procedure \cite{Legeza2003,fiedler,interface} as invoked by \texttt{dmrginit.py}.\\
Example: \texttt{hf\_occ = "4,4,4,2,2,1"}\ for a CAS(8,6) triplet state setup where the unpaired electrons are ``located"
in orbital \# 4 and \# 5.}\\
& \\[\tableoptionskip]
\multirow{2}{*}{\texttt{nmainsweeps}} & \multirow{2}{11cm}{Number of sweeps in which \texttt{truncation\_final} is used in the singular value decomposition.} \\
& \\[\tableoptionskip]
\multirow{2}{*}{\texttt{ngrowsweeps}} &\multirow{2}{11cm}{Number of sweeps in which \texttt{truncation\_initial} is used in the singular value decomposition.}\\
& \\[\tableoptionskip]
\texttt{storagedir} &Scratch directory for temporary files.\\

\subsubsection{Keywords for expectation value calculations}
\label{sec:prop-qcm}

\qcm\ can (in principle) compute expectation values for any one- or two-particle operator that can be formulated in
second quantization. Table \ref{tab:expval}\ comprises a list of the available property keywords available in \qcm.
% For further updates/other properties please contact \href{mailto:dmrg@phys.chem.ethz.ch}{dmrg@phys.chem.ethz.ch}.

The one-particle reduced density matrix as well as the one-particle spin-density matrix are implicitly computed from the
expectation values of some of the operators contained in \texttt{MEASURE[ChemEntropy]}.

If a property is available only in 2U1 symmetry, the SU2U1 checkpoint may be transformed into 2U1 group with the \texttt{mps\_transform(\_pg)} tool (see Section~\ref{sec:tools-qcm})

\begin{longtable}{ll}
 \caption{\label{tab:expval}Expectation value calculations available in the release version of \qcm.}
\\
  \toprule
keyword & description \\
  \midrule
\endfirsthead

\multicolumn{2}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\toprule
keyword & description \\
\midrule
\endhead
% \begin{tabular}{p{5.6cm}l}

\multirow{3}{*}{\texttt{MEASURE[ChemEntropy]}} & \multirow{3}{11cm}{All expectation values over the operators required to
calculate the mutual information (as specified in Ref.~\cite{entanglement}) will be computed. Available only for SU2 symmetry.} \\
 & \\
 & \\[\tableoptionskip]
\multirow{3}{*}{\texttt{MEASURE[1rdm]}} & \multirow{3}{11cm}{Computes the one-particle reduced density matrix, without the additional correlators
contained in the \texttt{ChemEntropy} measurement.}\\
 & \\
 & \\[\tableoptionskip]
\multirow{10}{*}{\begin{minipage}{5cm}
                 \texttt{MEASURE[2rdm]}\\
                 \texttt{MEASURE[3rdm]}\\
                 \texttt{MEASURE[4rdm]}\\
                \end{minipage}
                }& \multirow{10}{11cm}{Computes the two-particle, three-particle and four-particle reduced density matrix, respectively. In case of 4-RDM, \texttt{MEASURE[4rdm]="p4:p3:p1:p2@\textit{l,k,i,j}"} allows to compute only a 4-index sub-block of the 4-RDM with the indices \texttt{\textit{i,j,k,l}}. This functionality is used in the distributed RDM calculations described in Section~\ref{sec:4rdm-distributed}. Analogously, for 3-RDM a two-index subblock may be calculated with \texttt{MEASURE[3rdm]="p1:p2@\textit{j,i}"}.\\ 3-RDM and 4-RDM calculations are available only in 2U1(pg) symmetry.}\\
& \\
& \\
&\\
&\\
&\\
&\\
&\\
&\\
&\\[\tableoptionskip]
\multirow{6}{*}{\begin{minipage}{5.6cm}
%                  \texttt{MEASURE[trans1rdm]="\textit{chk}"}\\ %% trans1rdm_aa, _bb, _ab, _ba ??
                 \texttt{MEASURE[trans2rdm]="\textit{chk}"}\\
                 \texttt{MEASURE[trans3rdm]="\textit{chk}"}\\
                \end{minipage}} & \multirow{6}{11cm}{Computes the transition two- or three-particle reduced density matrix with the current state and the state specified in the \texttt{\textit{chk}} checkpoint file. For t-3RDM calculations, two-index subblocks can be calculated with \texttt{MEASURE[trans3rdm]="\textit{chk};p1:p2@\textit{j,i}"}. Available only for 2U1(pg) symmetry.}\\
 & \\
 & \\
 & \\
 & \\
 & \\ \pagebreak
\multirow{4}{*}{\texttt{MEASURE\_LOCAL[\textit{name}]="\textit{op}"}} &
            \multirow{4}{11cm}{Computes $\langle \psi | op_i | \psi \rangle, \, i = 1 \ldots L$. \texttt{Nup}, \texttt{Ndown} and \texttt{Nup*Ndown}
            are meaningful choices for \textit{op}. Available for \texttt{2u1(pg) only.}\\
            \emph{Note}: \texttt{\textit{name}}\ defines the name under which the expectation values are stored on the \texttt{resultfile}.}\\
 & \\
 & \\
 & \\[\tableoptionskip]
\multirow{8}{5.6cm}{\texttt{MEASURE\_HALF}\texttt{\_CORRELATIONS} \texttt{[\textit{name}]="\textit{op$_1$:op$_2$}"}} &
            \multirow{8}{11cm}{Computes $\langle \psi | op_{1i} op_{2j}| \psi \rangle, \, i = 1 \ldots L, j=i+1 \ldots L$.
            \texttt{Nup}, \texttt{Ndown}, \texttt{Nup*Ndown}, \texttt{cdag\_up}, \texttt{cdag\_down}, \texttt{c\_up}, \texttt{c\_down},
            \texttt{cdag\_up*Ndown},
            \texttt{c\_up*Ndown},
            \texttt{cdag\_down*Nup},
            \texttt{c\_down*Nup},
            \texttt{cdag\_up*cdag\_down},
            \texttt{c\_up*c\_down},
            \texttt{cdag\_up*c\_down},
            and \texttt{cdag\_down*c\_up},
            as \textit{op$_1$} and \textit{op$_2$} are recognized by the program. Available for \texttt{2u1(pg)} only.\\
            \emph{Note}: \texttt{\textit{name}}\ defines the name under which the expectation values are stored on the \texttt{resultfile}.}\\
%\multirow{2}{*}{= "\textit{op$_1$:op$_2$}"}} & \\
 & \\
 & \\
 & \\
 & \\
 & \\
 & \\
 & \\
\bottomrule
\end{longtable}
\FloatBarrier

\subsection{QCMaquis input example}
\label{sec:input-qcm}

Listing~\ref{lst:qcm} provides an input example for a DMRG-CASCI calculation.
It assumes that an integral file \texttt{FCIDUMP} has been generated and placed in the same directory with the input file.
The mandatory keywords are indicated in red, and the optional keywords in blue.
The \texttt{FCIDUMP} for this particular example has been generated by the \mol-\qcm{} interface with the \texttt{001.input} input file from the \mol-\qcm{} test directory.

\begin{lstlisting}[language=qcmaquis,caption={Input example for a DMRG-CASCI(8,8) standalone \qcm{} calculation},label=lst:qcm]
  // Lattice size (= number of orbitals)
  L     = 8
  // Number of electrons
  nelec = 8
  // Target symmetry (spin and spatial) of the wave function:
  // 2*spin = 0 (singlet)
  spin  = 0
  // Wavefunction irreducible representation (0 == totally symmetric)
  irrep = 0
  // Parameters to control the actual DMRG calculation
  nsweeps              = 8
  max_bond_dimension   = 256
  conv_thresh          = 1e-6
  // MPS initialization
  init_state = 'default'
  // Wave function symmetry (su2u1pg for spin-adapted DMRG + point group)
  symmetry             = 'su2u1pg'
  // Threshold below which integrals inside the integral_file are neglected
  integral_cutoff      = 1e-40
  // Threshold for the SVD truncation for the initial sweeps
  truncation_initial   = 1e-16
  // Threshold for the SVD truncation for the final sweeps
  truncation_final     = 1e-7
  // HDF5 file where the MPS will be stored
  chkpfile             = 'chkp.h5'
  // HDF5 file where the results of the DMRG calculations will be stored
  resultfile           = 'result.h5'
  // File with the one- and two-body integrals
  integral_file        = "FCIDUMP"
  // Path to the scratch folder
  storagedir           = '/scratch/$USER/boundaries'
  // Lattice-related parameters (not to be changed for qc calculations)
  LATTICE              = "orbitals"
  lattice_library      = "coded"
  MODEL                = "quantum_chemistry"
  model_library        = "coded"
  // List of measurements to be performed at the end of the optimization
  MEASURE[ChemEntropy] = 1
\end{lstlisting}

\section{QCMaquis vibrational DMRG module}
\label{sec:vDMRG}

The vibrational DMRG (vDMRG) module enables calculating ground- and excited-state energies of the Born--Oppenheimer vibrational Hamiltonian:

\begin{equation}
  \underbrace{\left[ \mathcal{T}(Q_1, \ldots, Q_L) + \mathcal{V}(Q_1, \ldots, Q_L) \right]}_{\mathcal{H}_\text{vib}(Q_1, \ldots, Q_L)} \Psi_n(Q_1, \ldots, Q_L) = E_n \Psi_n(Q_1, \ldots, Q_L)
  \label{eq:VibHam}
\end{equation}

We designed two vDMRG formulations, one based on the so-called canonical quantization format,\cite{Hirata2014_NormalOrdered} and one based on the $n$-mode second quantization.\cite{Christiansen2004_nModeQuantization}
These two formulations differ in the way in which the Hamiltonian $\mathcal{H}_\text{vib}$ is encoded based on the second-quantization formalism.
In the following, we will describe how this encoding is obtained.

\subsection{Canonical quantization vDMRG}
\label{sec:canonical}

The canonical quantization relies on the harmonic oscillator theory.\cite{Hirata2014_NormalOrdered}
The vibrational wave function is expanded, in practice, as follows:

\begin{equation}
  \Psi(Q_1, \ldots, Q_L) = \sum_{n_1}^{N_\text{max}} \cdots \sum_{n_L=1}^{N_\text{max}}
	C_{n_1 \cdots n_L} | n_1 \cdots n_L \rangle \, .
  \label{eq:CanonicalQuantizationCI}
\end{equation}

$| n_1 \cdots n_L \rangle$ are occupation number vectors that are constructed using the harmonic oscillator eigenfunction as the one-particle basis set, \textit{i.e.}

\begin{equation}
  | n_1 \cdots n_L \rangle \equiv \prod_{i=1}^L \chi_{n_i}(Q_i)
  \label{eq:HarmonicOscillatorBasis}
\end{equation}
%
$\chi_{n_i}(Q_i)$ being the $n_i$-th eigenfunction of the monodimensional harmonic oscillator Hamiltonian associated with the $i$-th mode.
In order to encode the Hamiltonian in a format analog to Eq.~\ref{eq:CanonicalQuantizationCI}, we express it as a power series in terms of the vibrational coordinates $Q_i$.
By including up to fourth-order force constants, the expansion reads:

\begin{equation}
  \mathcal{V}(Q_1, \ldots, Q_L) = \sum_{i,j=1}^L k_{ij} Q_i Q_j + \sum_{i,j,k=1}^L k_{ijk} Q_i Q_j Q_k
    + \sum_{i,j,k,l=1}^L k_{ijkl} Q_i Q_j Q_k Q_l
  \label{eq:TaylorExpansion}
\end{equation}
%
where we neglected the first-order force constants since we assume that the expansion is performed around a stationary point of the PES.
The position operators $Q_i$ are, then, encoded in second quantization based on the well-known canonical quantization rules.

\begin{equation}
  Q_i = \frac{1}{2} \left( b_i^\dagger + b_i \right) \, .
  \label{eq:CanonicalQuantizationPosition}
\end{equation}

Note that Eq.~\ref{eq:CanonicalQuantizationPosition} holds only if $Q_i$ are the dimensionless Cartesian-based normal coordinates.
Moreover, the kinetic energy operator can be expressed in second quantization based on the counterpart of Eq.~(\ref{eq:CanonicalQuantizationPosition}) for the conjugate momentum operator

\begin{equation}
  P_i = \frac{\mathrm{i}}{2} \left( b_i^\dagger - b_i \right) \, .
  \label{eq:CanonicalQuantizationMomentum}
\end{equation}

The resulting form of the vibrational Hamiltonian is then converted in the MPO format by \qcm\ based on the algorithm described in Ref.~\citenum{Baiardi2017_vDMRG}. \\

\begin{lstlisting}[language=qcmaquis,
				   caption={Input example for a canonicalq quantization vDMRG-FCI calculation},
				   label=lst:canonicalVibInput]
  // == Generic input section ==
  nsweeps                         = 6
  max_bond_dimension              = 100
  model_library                   = coded
  lattice_library                 = coded
  optimization                    = twosite 
  integral_file                   = FCIDUMP_Vibrational
  integral_cutoff                 = 1.0E-10
  init_state                      = 'default'
  L                               = 12
  // == vDMRG-specific input ==
  symmetry                        = none
  LATTICE                         = 'watson lattice'
  MODEL                           = 'watson'
  Nmax                            = 10
\end{lstlisting}

We report in Listing~\ref{lst:canonicalVibInput} an example input for a vDMRG calculation relying on the canonical-quantization based formalism.
The first part of the input has the same structure as that of an electronic-structure DMRG calculation.
Note that the lattice dimension $L$ is, in this case, equal to the overall number of modes of the molecule.
Unlike for the electronic-structure case, the DMRG symmetry must be set to \texttt{none}, and the \texttt{MODEL} keyword should be set to \texttt{watson}.
As the expansion reported in Eq.~\ref{eq:CanonicalQuantizationCI} is virtually unbounded, it must be truncated, in practice, by setting a threshold $N_\text{max}$ for the maximum excitation degree of a mode.
Such a threshold must be specified in input, as shown in Listing~\ref{lst:canonicalVibInput}.

The integral file \texttt{FCIDUMP\_Vibrational} contain the force constants that define the PES through Eq.~\ref{eq:TaylorExpansion}.
\qcm\ supports up to sixth-order force constants, that must be provided as input in the \texttt{FCIDUMP\_Vibrational} file as illustrated in Listing~\ref{lst:CanonicalFCIDUMP}
\vspace{.5cm}
\begin{lstlisting}[language=qcmaquis,caption={Example of an vDMRG integral file for the potential
											  expressed in the canonical quantization format.},
				   label=lst:CanonicalFCIDUMP, mathescape=true, rulecolor=\color{black}]
  $\smash{k_{ij}}$     i    j   0   0   0   0
  $\smash{\vdots}$
  $\smash{h_{ij}}$    -i   -j   0   0   0   0
  $\smash{\vdots}$
  $\smash{k_{ijk}}$     i    j   k   0   0   0
  $\smash{\vdots}$
  $\smash{k_{ijkh}}$     i    j   k   h   0   0
  $\smash{\vdots}$
  $\smash{k_{ijkhm}}$     i    j   k   h   m   0
  $\smash{\vdots}$
  $\smash{k_{ijkhmn}}$     i    j   k   h   m   n
\end{lstlisting}
\vspace{.5cm}
The $h$ and $k$ constants define the PES as follows:

\begin{equation}
  \begin{aligned}
    V(\tilde{Q}_1, \ldots, \tilde{Q}_L) 
      &= \sum_{i=1}^L k_{ii} \tilde{Q}_i^2 + \sum_{i=1}^L h_{ii} \tilde{P}_i^2 
      + \sum_{ijk=1}^L k_{ijk} \tilde{Q}_i \tilde{Q}_j \tilde{Q}_k
      + \sum_{ijkl=1}^L k_{ijkl} \tilde{Q}_i \tilde{Q}_j \tilde{Q}_k \tilde{Q}_l \\
      &+ \sum_{ijklm=1}^L k_{ijkl} \tilde{Q}_i \tilde{Q}_j \tilde{Q}_k \tilde{Q}_l \tilde{Q}_m
      + \sum_{ijklmn=1}^L k_{ijkl} \tilde{Q}_i \tilde{Q}_j \tilde{Q}_k \tilde{Q}_l \tilde{Q}_m \tilde{Q}_n \, ,
  \end{aligned}
  \label{eq:PES_Representation}
\end{equation}
%
where we introduced the scaled dimensionless normal coordinates $\tilde{Q}_i$ that are defined as:

\begin{equation}
  \tilde{Q}_i = \sqrt{2} Q_i = \left( b_i^\dagger + b_i \right)
  \label{eq:ScaledNM}
\end{equation}
%
$Q_i$ being the $i$-th dimensionless normal coordinate operator.
The corresponding scaled dimensionless momentum operator $\tilde{P}_i$ is defined as:

\begin{equation}
  \tilde{P}_i = \frac{\sqrt{2}}{\mathrm{i}} P_i = \left( b_i^\dagger - b_i \right)
  \label{eq:ScaledNM_Momentum}
\end{equation}

\subsection{$n$-mode vDMRG}
\label{sec:nMode}

The canonical quantization-based vDMRG theory requires the PES to be expressed as a Taylor series in terms of the Cartesian-based normal coordinates.
Such a format is, however, ill-suited to describe
\qcm\ can apply the vDMRG method to potentials that are encoded as generic many-body expansions \cite{Bowman1997_ManyBody}

>>>>

% -- Loads packages --
\usepackage[nott]{kpfonts}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{braket}
\usepackage[bf,singlelinecheck=off]{caption}
\usepackage[T1]{fontenc}
\usepackage{framed}
\usepackage{graphicx}
\usepackage[colorlinks=true]{hyperref} % web links
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{longtable}
\usepackage{multirow}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{placeins}
\usepackage{xcolor}

@@

\lstdefinelanguage{qcmaquis}
{
%
    keywords=[1]{L,nelec,spin,irrep,nsweeps,max_bond_dimension,integral_file,chkpfile,resultfile,init_type,LATTICE,lattice_library,MODEL,model_library},
	morekeywords=[2]{conv_thresh,integral_cutoff,truncation_initial,truncation_final,symmetry,MEASURE},
	keywordstyle=[1]\color{green!20!red},
	keywordstyle=[2]\color{green!20!blue},
	commentstyle=\color{gray},
	sensitive=false,
	morecomment=[l]{//},
	morecomment=[s]{/*}{*/},
}

@@
